---
title: "Simple data analysis on Weight Lifting Exercise Dataset"
author: "Lu Chai"
date: "02/21/2015"
output: html_document
---

Synopsis
---------------------------------------------------
In this report we first read in the provided training and testing dataset. Next, we clean up the raw data: 1) we remove columns that only contains #DIV/0! and blank and rows with #DIV/0! as it is not valid for the numeric data type. 2) We convert factors to numerics. 3) We drop zero columns from a near zero analyses. Then, we create a model using random forest. Lastly, we apply the model to our testing set to predict our outcome.

# Data Processing

```{r}
library(caret)
# function to clean up our raw data
preprocessData <- function(df) {
  # remove columns with #DIV/0!
  df <- df[,-which(names(df) %in% c("new_window","amplitude_yaw_belt", "amplitude_yaw_forearm","amplitude_yaw_dumbbell", "kurtosis_yaw_forearm", "kurtosis_yaw_belt", "skewness_yaw_belt", "kurtosis_yaw_dumbbell", "skewness_yaw_dumbbell", "skewness_yaw_forearm"))]
  
  dropInd <- which(apply(df, 1, function(x) any(x=="#DIV/0!")))
  df <- df[-dropInd,]
  
  library(caret)
  
  # convert factors to numeric 
  facs <- sapply(df[,-ncol(df)], is.factor)
  df[ , facs] <- as.numeric(as.character(df[ , facs]))
  
  # set NA to 0
  df[is.na(df)] <- 0
  
  nsv <- nearZeroVar(df,saveMetrics=TRUE)
  
  # remove zero columns
  df <- df[,-which(names(df) %in% c(rownames(nsv[nsv$zeroVar,])))]
  return(df)
}

# set seed
set.seed(2172015)

# load training and testing dataset
training <- read.csv("pml-training.csv")
testing <- read.csv("pml-testing.csv")

# clean up training data using the function we created 
# also apply same processing to the testing dataset
training<-preprocessData(training)
testing[is.na(testing)] <- 0
testing <- testing[, which(names(testing) %in% names(training))]

```

# Model creation

We use 2-fold cross-validation
```{r}
# create training control object for 2-fold cross-validation
fitControl <- trainControl(method="cv",
                           number=2,
                           repeats=1,
                           verboseIter=TRUE)
```

Random forest model

```{r}
# Random Forest
# with principle component analysis
rf_model <- train(classe~.,data=training,method="rf",
                 trControl=fitControl,
                 preProcess=c("scale", "pca")) 
rf_model

```
With this model, we achieved accuracy of 0.9786960 and standard deviation of 0.002184765. (See output above.)

# Result
Apply our random forest model to the testing dataset to obtain predictions
```{r}
testRf<- predict(rf_model, testing)
testRf
```